Information Aggregation in Exponential Family Markets

### 2014-02-12 WriteUp/maxentpredmkts.tex
we are at 18...

Scoring rules are a measure of prediction accuracy, and we are concerned here with scoring rules for statistic expectations, typically over infinite outcome spaces. Such rules have been characterized by~\citet{Savage71}; see also~\citep{frongillo2013elicitingmeans,lambert2008eliciting}. Our rules are of course special cases of this characterization, but it appears the range of elegant scoring rules that arise from exponential families has not been appreciated. Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities. In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.

@@

Prediction markets have been known to perform well in practice~\citep{AGT07,Wol04,pennock2001real}. However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study~\citep{wolfers2006interpreting}. At one extreme, agents are assumed myopic and risk-neutral, which implies that they move the market state to their belief~\citep{chen2010new}. At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}. Strategic behavior in prediction markets is also addressed in~\citep{chen2010gaming,dimitrov08}.

@@

We are interested in eliciting information about the agent's belief, in particular expectation information. Let $\phi: \mX \rightarrow \bR^d$ be a vector-valued random variable or \emph{statistic}, where $d$ is finite. The aim is to elicit $\mu = \E_{p}[\phi(x)]$ where $x$ is the random outcome. A \emph{scoring rule} is a device for this purpose. Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations.  A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$. The following definition is due to~\citet{lambert2008eliciting}.
%
\begin{definition} \label{def:proper-score}
A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have
%
\begin{equation} \label{eq-proper}
\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)]
\end{equation}
%
for all alternative $\hmu \neq \mu$.% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
\end{definition}
%

@@

When the outcome space is finite we can recover classical scoring rules from the definition by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$. The expectation of $\phi$ is then exactly the probability mass function.

@@

The following observation is crucial. Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report. Then from~(\ref{exp-fam}) note
%
\begin{equation} \label{equalizer-rule}
\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
\end{equation}
%
where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$. We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%
\begin{theorem} \label{maxent-score}
Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$. The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$. 
\end{theorem}
%
%
\begin{proof}
Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$. Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$. We have
%
\begin{eqnarray}
& & \Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)] \nonumber\\
& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
\end{eqnarray}
%
The latter is positive by the strict convexity of $T$, which shows that the log score is proper. For the converse, assume the defined log score is proper. By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{gneiting2007strictly} and~\citet{Savage71}), we must have
%
$$
S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
$$
for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
%
\begin{eqnarray*}
\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra \\
& = & \la \theta, \phi(x) \ra - T(\theta),
\end{eqnarray*}
%
which shows that $p(x;\mu)$ takes the form of an exponential family.
\end{proof}
%

@@

An advantage of this generalization of the log score is that, for many domains (multi-dimensional included) and expectations of interest, it leads to novel closed-form scoring rules. By examining the log densities of various exponential families, we can for instance obtain scoring rules for several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments. The following examples illustrate the construction.
%
\begin{example} \label{ex:exponential}
As base measure we take the Lebesgue restricted to $[0,+\infty)$, and we consider the statistic $\phi(x) = x$ so that we are simply eliciting the mean. The maximum entropy distribution with a given mean $\mu$ is the exponential distribution, and taking its log density gives the scoring rule
%
\begin{equation} \label{exp-scoring}
S(\mu, x) = -\frac{x}{\mu} - \log \mu.
\end{equation}
%
We stress that although this rule is derived from the exponential distribution, Theorem~\ref{maxent-score} implies that it elicits the mean of any distribution supported on the non-negative reals (e.g., Pareto, lognormal). Indeed, it is easy to see that the expected score~(\ref{exp-scoring}) depends only on the mean of the agent's belief because it is linear in $x$. 
%
%Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$. In that case the rule takes the form $S(\mu, x) = -\lambda x + \log \lambda$, and we see that $\lambda$ corresponds to the natural parameter.
%
As a generalization of this example, the maximum entropy distribution for the $k$-th moment $\phi(x) = x^k$ with respect to the same base measure is the Weibull distribution. Taking its log density leads to the following equivalent scoring rule:
%
\begin{equation} \label{weibull-scoring}
S(\mu, x) = (k-1) \log x - k \log \mu - \Gamma\left(1+\frac{1}{k}\right)^k \left(\frac{x}{\mu}\right)^k,
\end{equation}
%
where $\Gamma$ denotes the gamma function (the extension of the factorial to the reals). We have not found either scoring rule~(\ref{exp-scoring}) or~(\ref{weibull-scoring}) in the literature.
\end{example}
%
%
\begin{example} \label{ex:gaussian}
As a base measure we take the Lebesgue over the real numbers $\bR$. We are interested in eliciting the mean $\mu$ and variance $\sigma^2$, so as a statistic we take $\phi(x) = (x, x^2)$ for which $\Exp_p[\phi(x)] = (\mu, \mu^2+\sigma^2)$. The maximum entropy distribution for a given mean and variance is the normal distribution, and taking its log density gives the scoring rule
%
\begin{equation}
S((\mu, \sigma^2), x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2.
\end{equation}
%
Again, we stress that this scoring rule elicits the mean and variance of any density over the real numbers, not just those of a normal distribution. The construction easily generalizes to a multi-dimensional outcome space by taking the log density of the multivariate normal:
%
\begin{equation}
S((\mu, \Sigma), x) = -(x - \mu)'\Sigma^{-1}(x - \mu) - \log |\Sigma|.
\end{equation}
%
Here the statistics being elicited are the mean vector $\mu$ and the covariance matrix $\Sigma$. These scoring rules have been studied by~\citet{dawid1999coherent} as rules that only depend on the mean and variance of the reported density. They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
\end{example}
%

@@

In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{chen2007utility,chen2010new}. As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$. However, this means that the market does not perform any meaningful aggregation of the agent's belief, as the final prices are simply the final agent's expectation.

@@

In this section we relax the standard assumption that agents in the market are risk-neutral. We show that, with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' expectations in the form of a simple weighted average.
Assume that the agent has an exponential utility function for wealth $w$:
%
\begin{equation} \label{exp-util}
U_a(w) = -\frac{1}{a} \exp(-aw).
\end{equation}
%
Here $a$ controls the level of risk aversion: the agent is more risk averse as $a$ increases, and as $a$ tends to 0 we approach linear utility (risk-neutrality). Specifically, $a$ is the Arrow-Pratt coefficient of absolute risk aversion, and exponential utilities of the form~(\ref{exp-util}) are the unique utilities that exhibit constant absolute risk aversion~\citep[Chap.\ 11]{VarianMicro}.

@@

\noindent
Recall that for the generalized LMSR, the cost function $C$ is exactly the log partition function $T$. We are therefore lead to the following understanding of a risk-averse agent's behavior in such a market.
%
\begin{theorem} \label{thm:exp-util}
Suppose an agent has exponential utility with coefficient $a$ and exponential family beliefs with natural parameter $\htheta$. In the generalized LMSR market with current market state $\theta$, the agent's optimal trade $\delta$ moves the state vector to
%
\begin{equation} \label{exp-update}
\theta + \delta = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta.
\end{equation}
%
\end{theorem}
%
%
\begin{proof}
The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and the fact that $T = C$, the agent therefore maximizes
%
$$
\log a - T(\htheta - a\delta) + T(\htheta) - aT(\theta+\delta) + aT(\theta). 
$$
This objective is strictly concave, from the strict convexity of $T$. The optimum is therefore characterized by the first-order condition $\grad T(\htheta - a\delta) = \grad T(\theta+\delta)$. As the gradient map $\grad T$ is one-to-one, this is solved by equating the arguments, which leads to $\delta = (\htheta - \theta) / (1+a)$ and~(\ref{exp-update}).
% 
\end{proof}
%

@@

In the context of the generalized LMSR we write $T$ rather than $C$ where $T$ is the log partition function, with liquidity-adjusted version $T_{\lambda}$. Let $\hmu$ be the agent's mean belief with corresponding natural parameter $\htheta = \grad T^{-1}(\hmu)$. Recall that a risk-neutral agent moves the share vector so that the prices match its mean parameter. Therefore, define the \emph{target shares} as $\ttheta = \grad T^{-1}_{\lambda}(\hmu)$. The target shares $\htheta$ and natural parameter $\ttheta$ are related by
%
$$
\grad T(\htheta) = \grad T_{\lambda}(\ttheta) = \hmu.
$$
%
It is straightforward to check that $\grad T_{\lambda}(\ttheta) = \grad T(\lambda \ttheta)$ so we have
%
\begin{equation} \label{target}
\ttheta = \htheta / \lambda.
\end{equation}
%
Higher price responsiveness means fewer shares must be bought to make the market prices match the agent's expectation, so the natural parameter is scaled down accordingly. With a liquidity adjustment the analysis of Theorem~\ref{thm:exp-util} can be extended and yields the following result, where as before $\theta$ and $\mu$ are the market's outstanding shares and prices respectively.
%
\begin{corollary}
Under the conditions of Theorem~\ref{thm:exp-util} and an inverse liquidity of $\lambda$, the agent's optimal trade $\delta$ moves the state vector to
%
\begin{equation} \label{liqu-update}
\theta + \delta \:\:=\:\: \frac{\lambda}{\lambda+a} \ttheta + \frac{a}{\lambda+a} \theta
\:\:=\:\: \frac{\lambda}{\lambda+a} \grad T^{-1}_{\lambda}(\hmu) + \frac{a}{\lambda+a} \grad T^{-1}_{\lambda}(\mu).
\end{equation}
%
\end{corollary}
%
According to~(\ref{liqu-update}), as $\lambda$ grows large the agent moves the market state closer to the \emph{target shares}, rather than its true natural parameter. Note that the target shares themselves depend on $\lambda$ by~(\ref{target}), but the update can be directly written in terms of the agent's beliefs as in the right-hand side of~(\ref{liqu-update}).

@@

We draw a well-known result from game theory regarding the class of \emph{potential games}. We say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
%
\begin{theorem}[\citet{monderer1996potential}]\label{thm:potl}
Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}= (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$. If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
\end{theorem}
%
In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).  Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market when the market has reached equilibrium. Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his risk aversion parameter.
% \jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}

@@

Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is:
$$\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$$
\end{proof}

>>>>

Scoring rules are a measure of prediction accuracy, and we are concerned here with scoring rules for statistic expectations, typically over infinite outcome spaces. Such rules have been characterized by~\citet{Savage71}; see also~\citep{frongillo2013elicitingmeans,lambert2008eliciting}. Our rules are of course special cases of this characterization, but it appears the range of elegant scoring rules that arise from exponential families has not been appreciated. Indeed,~\citet{gneiting2007strictly} observe that specific instances of scoring rules for continuous outcomes are lacking, and survey various possibilities.
% In an early work,~\citet{matheson1976scoring} designed scoring rules for real-valued outcomes from those for binary outcomes.

@@

Prediction markets are known to perform well in practice~\citep{AGT07,pennock2001real}. However, a sound theory for interpreting trader behavior and market prices is an ongoing area of study~\citep{wolfers2006interpreting}. At one extreme, agents are assumed myopic and risk-neutral, implying they move the market state to their belief~\citep{chen2010new}. At the other extreme, agents are strategic and the market fully incorporates all information~\citep{ostrovsky2012}. Strategic behavior in prediction markets is addressed in~\citep{chen2010gaming,dimitrov08}.

@@

We are interested in eliciting information about the agent's belief, in particular expectation information. Let $\phi: \mX \rightarrow \bR^d$ be a vector-valued random variable or \emph{statistic}, where $d$ is finite. The aim is to elicit $\mu = \E_{p}[\phi(x)]$ where $x$ is the random outcome. A \emph{scoring rule} is a device for this purpose. Let $$\mM = \left\{ \mu \in \bR^d : \Exp_p[\phi(x)] = \mu,\, \mbox{for some $p \in \mP$} \right\}$$ be the set of realizable statistic expectations.  A scoring rule $S : \mM \times \mX \rightarrow \bR \cup \{-\infty\}$ pays the agent $S(\hmu,x)$ according to how well its report $\hmu \in \mM$ agrees with the eventual outcome $x \in \mX$. The following definition is due to~\citet{lambert2008eliciting}.
%
\begin{definition} \label{def:proper-score}
A scoring rule $S$ is  \emph{proper} for statistic $\phi$ if for each $\mu \in \mM$ and $p \in \mP$ with expected statistic $\mu$, we have for all $\hmu \neq \mu$
%
\begin{equation} \label{eq-proper}
\Exp_p[S(\mu,x)] \geq \Exp_p[S(\hmu,x)].
\end{equation}
%
% If the domain is $\mD = \mP$, the set of all possible densities, then we simply say the scoring rule is \emph{proper}.
\end{definition}
%

@@

When the outcome space is finite we recover classical scoring rules by using the statistic $\phi : \mX \rightarrow \{0,1\}^{\mX}$ that maps an outcome $x$ to a unit vector with a 1 in the component corresponding to $x$. The expectation of $\phi$ is then exactly the probability mass function.

@@

The following observation is crucial. Let $\tp \in \mP$ be a density (not necessarily from an exponential family) with expected statistic $\mu$, let $p(\cdot\,;\mu)$ be the exponential family with the same expected statistic, and let $\hmu \in \mM$ be an alternative report. Then from~(\ref{exp-fam}) note
%
\begin{equation} \label{equalizer-rule}
\Exp_{\tp}[\log p(x;\hmu)] = \Exp_{p(\cdot;\mu)}[\log p(x;\hmu)] = \la \htheta, \mu \ra - T(\htheta),
\end{equation}
%
where $\htheta = \grad G(\hmu)$ is the natural parameter for the exponential family with statistic $\hmu$. We see from this that the expected log score only depends on the expectation $\mu$ of the underlying density, not the full density, which is how we can achieve proper scoring according to Definition~{\ref{def:proper-score}.
%
\begin{theorem} \label{maxent-score}
Consider the logarithmic scoring rule $S(\mu,x) = \log p(x;\mu)$ defined over a set of densities $\mD$ parametrized by $\mM$. The scoring rule is proper if and only if $\mD$ is the exponential family with statistic $\phi$. 
\end{theorem}
%
%
\begin{proof}
Let $\mu,\hmu \in \mM$ be the agent's true belief and an alternative report, and let $p \in \mP$ be a density consistent with $\mu$. Let $\theta = \grad G(\mu)$ and $\htheta = \grad G(\hmu)$, and note that $\mu = \grad T(\theta)$. We have
%
\begin{eqnarray}
\Exp_p[\log p(x;\mu)] - \Exp_p[\log p(x;\hmu)]
& = & \la \theta, \mu \ra - T(\theta) - \la \htheta, \mu \ra + T(\htheta) \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \mu \ra \nonumber\\
& = & T(\htheta) - T(\theta) - \la \htheta - \theta, \grad T(\theta) \ra. \label{breg-div}
\end{eqnarray}
%
The latter is positive by the strict convexity of $T$, which shows that the log score is proper. For the converse, assume the defined log score is proper. By the Savage characterization of proper scoring rules for expectations (see~\citet[Thm.\ 1]{gneiting2007strictly} and~\citet{Savage71}), we must have
%
$$
S(\mu,x) = G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra
$$
for some strictly convex function $G$. Let $T = G^*$, so that $\grad G = {\grad T}^{-1}$, and let $\theta = \grad G(\mu)$. Then the above can be written as
%
\begin{eqnarray*}
\log p(x;\mu) & = & G(\mu) - \la \grad G(\mu), \mu - \phi(x) \ra \\
& = & \la \theta, \mu \ra - T(\theta) - \la \theta, \mu - \phi(x) \ra = \la \theta, \phi(x) \ra - T(\theta),
\end{eqnarray*}
%
which shows that $p(x;\mu)$ takes the form of an exponential family.
\end{proof}
%

@@

An advantage of this generalization of the log score is that, for many domains (multi-dimensional included) and expectations of interest, it leads to novel closed-form scoring rules. By examining the log densities of various exponential families, we can for instance obtain scoring rules for several different combinations of the arithmetic, geometric, and harmonic means, as well as higher order moments. The following examples illustrate the construction.
%
\begin{example} \label{ex:exponential}
As base measure we take the Lebesgue restricted to $[0,+\infty)$, and we consider the statistic $\phi(x) = x$ so that we are simply eliciting the mean. The maximum entropy distribution with a given mean $\mu$ is the exponential distribution, and taking its log density gives the scoring rule
%
\begin{equation} \label{exp-scoring}
S(\mu, x) = -\frac{x}{\mu} - \log \mu.
\end{equation}
%
We stress that although this rule is derived from the exponential distribution, Theorem~\ref{maxent-score} implies that it elicits the mean of any distribution supported on the non-negative reals (e.g., Pareto, lognormal). Indeed, it is easy to see that the expected score~(\ref{exp-scoring}) depends only on the mean of the agent's belief because it is linear in $x$. 
%
%Typically the exponential distribution is parametrized by its rate $\lambda = 1/\mu$. In that case the rule takes the form $S(\mu, x) = -\lambda x + \log \lambda$, and we see that $\lambda$ corresponds to the natural parameter.
%
As a generalization of this example, the maximum entropy distribution for the $k$-th moment $\phi(x) = x^k$ with respect to the same base measure is the Weibull distribution. Taking its log density leads to the following equivalent scoring rule:
%
\begin{equation} \label{weibull-scoring}
S(\mu, x) = (k-1) \log x - k \log \mu - \Gamma\left(1+\frac{1}{k}\right)^k \left(\frac{x}{\mu}\right)^k,
\end{equation}
%
where $\Gamma$ denotes the gamma function (the extension of the factorial to the reals). We have not found either scoring rule~(\ref{exp-scoring}) or~(\ref{weibull-scoring}) in the literature.
\end{example}
%
%
\begin{example} \label{ex:gaussian}
As a base measure we take the Lebesgue over the real numbers $\bR$. We are interested in eliciting the mean $\mu$ and variance $\sigma^2$, so as a statistic we take $\phi(x) = (x, x^2)$ for which $\Exp_p[\phi(x)] = (\mu, \mu^2+\sigma^2)$. The max entropy distribution for a given mean and variance is the Gaussian, whose log density gives the scoring rule
%
\begin{equation}
S((\mu, \sigma^2), x) = -\frac{(x-\mu)^2}{\sigma^2} - \log \sigma^2.
\end{equation}
%
Again, we stress that this scoring rule elicits the mean and variance of any density over the real numbers, not just those of a normal distribution. The construction easily generalizes to a multi-dimensional outcome space by taking the log density of the multivariate normal:
%
\begin{equation}
S((\mu, \Sigma), x) = -(x - \mu)'\Sigma^{-1}(x - \mu) - \log |\Sigma|.
\end{equation}
%
Here the statistics being elicited are the mean vector $\mu$ and the covariance matrix $\Sigma$. These scoring rules have been studied by~\citet{dawid1999coherent} as rules that only depend on the mean and variance of the reported density. They note that these rules are weakly proper (because they do not distinguish between densities with the same first and second moments), but do not make the point that knowledge of the full density is not necessary on the part of the agent.
\end{example}
%

@@

In the standard model of cost-function based prediction markets, a sequence of myopic, risk-neutral agents arrive and trade in the market~\citep{chen2007utility,chen2010new}. As we saw in Section~\ref{sec:pred-market}, such a trader moves the prices to its own expectation $\mu$. However, this means that the market does not perform meaningful aggregation of agents' beliefs, as the final prices are simply the final agent's expectation.

@@

In this section we relax the standard assumption that agents in the market are risk-neutral. We show that with sufficient extra structure to the agents' beliefs and utilities, the market performs a clean aggregation of the agents' beliefs via a simple weighted average.
Assume that the agent has an exponential utility function for wealth $w$:
%
\begin{equation} \label{exp-util}
U_a(w) = -\frac{1}{a} \exp(-aw).
\end{equation}
%
Here $a$ controls the risk aversion: the agent's aversion grows as $a$ increases, and as $a$ tends to 0 we approach linear utility (risk-neutrality). Specifically, $a$ is the Arrow-Pratt coefficient of absolute risk aversion, and exponential utilities of the form~(\ref{exp-util}) are the unique utilities that exhibit constant absolute risk aversion~\citep[Chap.\ 11]{VarianMicro}.

@@

\noindent
Recall that for the generalized LMSR, the cost function $C$ is exactly the log partition function $T$. We are therefore lead to the following understanding of a risk-averse agent's behavior in such a market.
%
\begin{theorem} \label{thm:exp-util}
Suppose an agent has exponential utility with coefficient $a$ and exponential family beliefs with natural parameter $\htheta$. In the generalized LMSR market with current market state $\theta$, the agent's optimal trade $\delta$ moves the state vector to
%
\begin{equation} \label{exp-update}
\theta + \delta = \frac{1}{1+a} \htheta + \frac{a}{1+a} \theta.
\end{equation}
%
\end{theorem}
%
%
\begin{proof}
The agent's optimal trade maximizes its expected utility, or equivalently the certainty equivalent. From Lemma~\ref{lem:exp-util} and that $T = C$, the agent maximizes
%
$$
\log a - T(\htheta - a\delta) + T(\htheta) - aT(\theta+\delta) + aT(\theta). 
$$
This objective is strictly concave, from the strict convexity of $T$. The optimum is therefore characterized by the first-order condition $\grad T(\htheta - a\delta) = \grad T(\theta+\delta)$. As the gradient map $\grad T$ is one-to-one, this is solved by equating the arguments, which leads to $\delta = (\htheta - \theta) / (1+a)$ and~(\ref{exp-update}).
% 
\end{proof}
%

@@

In the context of the generalized LMSR we write $T$ rather than $C$ where $T$ is the log partition function, with liquidity-adjusted version $T_{\lambda}$. Let $\hmu$ be the agent's mean belief with corresponding natural parameter $\htheta = \grad T^{-1}(\hmu)$. Recall that a risk-neutral agent moves the share vector so that the prices match its mean parameter. Therefore, define the \emph{target shares} as $\ttheta = \grad T^{-1}_{\lambda}(\hmu)$. The target shares $\htheta$ and natural parameter $\ttheta$ are related by $\grad T(\htheta) = \grad T_{\lambda}(\ttheta) = \hmu$. In addition it is straightforward to check that $\grad T_{\lambda}(\ttheta) = \grad T(\lambda \ttheta)$ so we have
%
\begin{equation} \label{target}
\ttheta = \htheta / \lambda.
\end{equation}
%
Higher price responsiveness means fewer shares must be bought to make the market prices match the agent's expectation, so the natural parameter is scaled down accordingly. With a liquidity adjustment the analysis of Theorem~\ref{thm:exp-util} can be extended and yields the following result, where as before $\theta$ and $\mu$ are the market's outstanding shares and prices respectively.
%
\begin{corollary}
Under the conditions of Theorem~\ref{thm:exp-util} and an inverse liquidity of $\lambda$, the agent's optimal trade $\delta$ moves the state vector to
%
\begin{equation} \label{liqu-update}
\theta + \delta \:\:=\:\: \frac{\lambda}{\lambda+a} \ttheta + \frac{a}{\lambda+a} \theta
\:\:=\:\: \frac{\lambda}{\lambda+a} \grad T^{-1}_{\lambda}(\hmu) + \frac{a}{\lambda+a} \grad T^{-1}_{\lambda}(\mu).
\end{equation}
%
\end{corollary}
%
According to~(\ref{liqu-update}), as $\lambda$ grows large the agent moves the market state closer to the \emph{target shares}, rather than its true natural parameter. Note that the target shares themselves depend on $\lambda$ by~(\ref{target}), but the update can be directly written in terms of the agent's beliefs as in the right-hand side of~(\ref{liqu-update}).

@@

We draw a well-known result from game theory regarding the class of \emph{potential games}. We say a function $f(\vec{x})$ is at a \emph{local optimum} if changing any coordinate of $\vec{x}$ does not increase the value of $f$.
%
\begin{theorem}[\citet{monderer1996potential}]\label{thm:potl}
Let $U_{i}(\vec{\delta})$ be the utility function of the $i^{th}$ trader given strategies $\vec{\delta}= (\delta_{1},\ldots,\delta_{i},\ldots,\delta_{n})$. If there exists a potential function $\Phi(\vec{\delta})$ such that $$U_{i}(\vec{\delta})-U_{i}(\vec{\delta}_{-i},\delta_{i}')=\Phi(\vec{\delta})-\Phi(\vec{\delta}_{-i},\delta_{i}')$$
then $\vec{\delta}$ is a Nash equilibrium if and only if $\Phi(\vec{\delta})$ is at a local optimum.
\end{theorem}
%
In the exponential family market, the cost function $C$ is identical to the log partition function $T$ defined in (\ref{log-part}).  Let $\vec{\delta}$ be the matrix of share vectors purchased by every trader in the market at equilibrium. Let $\theta$ be the initial market state, $\htheta_{i}$ the natural parameter of trader $i$'s belief distribution and $a_{i}$ his risk aversion parameter.
% \jake{I need to fix a small technical detail with the fact that we are using the log utilities to use the potential argument.}

@@

Since the market evolves so that the budget of any trader never falls below zero, the total myopic impact in $T$ rounds caused due to trader $i$ is $\Delta_{i}:=\sum_{t=1}^{T}\Delta_{i}^{t}= \sum_{t=1}^{T} (\alpha_{i}^{t} - \alpha_{i}^{t-1}) = \alpha_{i}^{T} - \alpha_{i}^{0}\geq -\alpha_{i}^{0}$.
\end{proof}
